The origin of cosmic-ray electrons in cluster outskirts

### 2012-01-26 paper.tex
update text

\subsection{CR electron distribution function} 
In a hot plasma the temperatures of electrons and protons (ions) are
equal. Under this constraint we derive the relationship between CR
electon momentum and CR proton momentum given by
%\begin{eqnarray}
%  \label{eq:pe_pp}
%\dd \Pe = \dd \Pp \frac{\sqrt{\Pe^2+2\left(m_\e c\right)^2-2m_\e m_\p\,c^2
%+2\left(m_\p-m_\e\right)c\sqrt{\left(m_\e c\right)^2+\Pe^2}}}{\Pe+\left(m_\p-m_\e\right) c \frac{\Pe}{\sqrt{\left(m_\e c\right)^2+\Pe^2}}}
%\end{eqnarray}
\begin{eqnarray}
  \label{eq:pe_pp}
\dd \pe &=& \dd \pp \,g(\pe) =  \dd \pp \,h(\pp)\,,\qquad \rmn{where}\\
g(\pe) &=& X \,\frac{\left[\pe^2-2\left(X-1\right)\left(1-\sqrt{1+\pe^2}\right)\right]^{0.5}}{\displaystyle\pe+\left(X-1\right)\frac{\pe}{\sqrt{1+\pe^2}}}\,,\qquad \rmn{and}\\
h(\pp) &=& X \,\frac{\displaystyle\pp-\left(1-Y\right)\frac{\pp}{\sqrt{1+\pp^2}}}{\left[\pp^2+2\left(1-Y\right)\left(1-\sqrt{1+\pp^2}\right)\right]^{0.5}}\,.\\
\end{eqnarray}
Note that $\pe = \Pe/m_\e c$ and $\pp = \Pp/m_\p c$, where we have
normalized the electron momentum $\Pe$ and the proton momentum $\Pp$
with the electron mass $m_\e$ and proton mass $m_\p$, respectively. We
have introduced the mass ratios $X=m_\p/m_\e$ and $Y=m_\e/m_\p$. Also
note that $\dd \Pe\rightarrow \dd \Pp$ in the asymptotic limit where
$\pp\gg 1$, and that $\dd \Pe\rightarrow \sqrt{m_\e / m_\p}\,\dd \Pp$ for
$\pp\ll 1$.

@@

\subsection{Synchrotron emission}
The process of merging clusters is often very violent where large
amounts of gravitational energy beeing dissipated in the form of
radiation, increased temperature, turbulence flows and shocks. The two
latter procesess are belived to accelerate CRs to high energy, where
especially the distribution function of preexisting CRs is
enhanced. In this section we use the smooth CR electron distribution
derived from a simulated merging cluster in the previous section,
boost it using the details of the shock, and then finally derive the
radio synchrotron emission.

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.32\columnwidth]{./figures/f_z.g72a.1.4Rv.a24.full.v20.eps}
  \includegraphics[width=0.32\columnwidth]{./figures/f_z.g1a.1.4Rv.a24.full.v20.eps}
  \includegraphics[width=0.32\columnwidth]{./figures/f_z.g676.1.4Rv.a24.full.v20.eps}
  \caption{NEEDS TO BE WRITTEN, WED}
\end{minipage}
\end{figure*}

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{./figures/CRspec.alphaCheck.eps}
  \includegraphics[width=0.49\columnwidth]{./figures/CRspec.snapCheck.eps}
  \caption{NEEDS TO BE WRITTEN, WED}
\end{minipage}
\end{figure*}

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{./figures/CRvolume.g72a.1.4Rv.a24.full.140.v20.eps}
  \includegraphics[width=0.49\columnwidth]{./figures/CRvolume.g1a.1.4Rv.a24.full.020.v20.eps}
  \caption{NEEDS TO BE WRITTEN, WED}
\end{minipage}
\end{figure*}

\begin{figure}
  \includegraphics[width=1.0\columnwidth]{./figures/CRespec.all.1.4Rv.a24.full.020.v20.eps}
  \caption{NEEDS TO BE WRITTEN, WED}
\end{figure}

\begin{figure}
  \includegraphics[width=1.0\columnwidth]{./figures/old/CRespecRadial.Radial2Rv.a24.140.v16.eps}
  \caption{replace figure with media all clusters, NEEDS TO BE WRITTEN, WED}
\end{figure}

\begin{figure}
  \includegraphics[width=1.0\columnwidth]{./figures/CRespec.g72a.1.4Rv.a24.full.140.v20.eps}
  \caption{NEEDS TO BE WRITTEN, WED\del{Cosmic ray electron spectra in
      cluster outskirts at redshift zero. We show the CR electron
      distribution function weighted with the electron Lorentz factor
      ($\gam_\e$) squared for a cluster with a recent merger for
      region between (1.3-1.5)$\rvir$. The solid lines show the median
      CR electron spectrum where both radiative and adiabatic losses
      are included. The dotted lines shows the 68 percentiles. The
      dashed lines show the CR electron spectrum without Coulomb and
      radiative cooling. Note that the Coulomb cooling at low energies
      is very inefficient because of the low electron densities in the
      cluster outskirts.}\label{fig:e_spec}}
\end{figure}

>>>>

\subsection{CR electron distribution function} 
In a hot plasma the temperatures of electrons and protons (ions) are
equal. Under this constraint we derive the relationship between CR
electron momentum and CR proton momentum given by
%\begin{eqnarray}
%  \label{eq:pe_pp}
%\dd \Pe = \dd \Pp \frac{\sqrt{\Pe^2+2\left(m_\e c\right)^2-2m_\e m_\p\,c^2
%+2\left(m_\p-m_\e\right)c\sqrt{\left(m_\e c\right)^2+\Pe^2}}}{\Pe+\left(m_\p-m_\e\right) c \frac{\Pe}{\sqrt{\left(m_\e c\right)^2+\Pe^2}}}
%\end{eqnarray}
\begin{eqnarray}
  \label{eq:pe_pp}
\dd \pe &=& \dd \pp \,g(\pe) =  \dd \pp \,h(\pp)\,,\qquad \rmn{where}\\
g(\pe) &=& X \,\frac{\left[\pe^2-2\left(X-1\right)\left(1-\sqrt{1+\pe^2}\right)\right]^{0.5}}{\displaystyle\pe+\left(X-1\right)\frac{\pe}{\sqrt{1+\pe^2}}}\,,\qquad \rmn{and}\\
h(\pp) &=& X \,\frac{\displaystyle\pp-\left(1-Y\right)\frac{\pp}{\sqrt{1+\pp^2}}}{\left[\pp^2+2\left(1-Y\right)\left(1-\sqrt{1+\pp^2}\right)\right]^{0.5}}\,.\\
\end{eqnarray}
Note that $\pe = \Pe/m_\e c$ and $\pp = \Pp/m_\p c$, where we have
normalized the electron momentum $\Pe$ and the proton momentum $\Pp$
with the electron mass $m_\e$ and proton mass $m_\p$, respectively. We
have introduced the mass ratios $X=m_\p/m_\e$ and $Y=m_\e/m_\p$. Also
note that $\dd \Pe\rightarrow \dd \Pp$ in the asymptotic limit where
$\pp\gg 1$, and that $\dd \Pe\rightarrow \sqrt{m_\e / m_\p}\,\dd \Pp$ for
$\pp\ll 1$.

@@

\subsection{Synchrotron emission}
The process of merging clusters is often very violent where large
amounts of gravitational energy being dissipated in the form of
radiation, increased temperature, turbulence flows and shocks. The two
latter processes are believed to accelerate CRs to high energy, where
especially the distribution function of preexisting CRs is
enhanced. In this section we use the smooth CR electron distribution
derived from a simulated merging cluster in the previous section,
boost it using the details of the shock, and then finally derive the
radio synchrotron emission.

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.32\columnwidth]{./figures/f_z.g72a.1.4Rv.a24.full.v20.eps}
  \includegraphics[width=0.32\columnwidth]{./figures/f_z.g1a.1.4Rv.a24.full.v20.eps}
  \includegraphics[width=0.32\columnwidth]{./figures/f_z.g676.1.4Rv.a24.full.v20.eps}
  \caption{Time evolution of the cosmic ray electron spectra in
    cluster outskirts. We show the total volume weighted CR electron
    distribution function for the particles that reside end up in
    region between (1.3-1.5)$\rvir$ at redshift zero; a cluster with a
    recent merger (g72a, left panel), large cooling flow cluster (g1a,
    middle panel), and a small cooling flow cluster (g676, right
    panel). The different line colors show the CR electron spectra at
    different look-back times, where the most recent spectrum in black
    is at 13.6 Gyrs. Notice the larger scatter in the merging cluster,
    and the relative small difference in the cooling flow
    clusters. The shape of the distribution function is very similar
    between different cooling flow clusters, and slightly more
    stochastic for merging clusters.\label{fig:e_spec_z}}
\end{minipage}
\end{figure*}

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{./figures/CRspec.alphaCheck.eps}
  \includegraphics[width=0.49\columnwidth]{./figures/CRspec.snapCheck.eps}
  \caption{Testing the robustness of the CR electron spectra. Left
    panel shows how different cuts in injected spectra impact the
    total volume weighted CR electron spectrum with full cooling; no
    spectral cut (red line), cut particles with $\alpha_\inj>2.2$
    (yellow line), cut particles with $\alpha_\inj>2.4$ (green line),
    cut particles with $\alpha_\inj>2.6$ (blue line). Note that in the
    main CR model we cut all particles with an $\alpha_\inj>2.4$,
    which is a factor four smaller at low energies than the spectrum
    without a cut. Right panel shows the effect of time-resolution on
    the CR electron spectra. The median CR spectrum is shown with full
    cooling that includes radiative, Coulomb, and adiabatic losses
    (dash-dotted line), the solid line represents the total volume
    weighted CR electron spectrum with full cooling. The three line
    colors corresponds to the same cluster, a large coma like merging
    cluster, but with different time resolution between snapshots in
    the relevant regime; red lines corresponds to 100 Myrs between
    snapshots, orange lines corresponds to ~(1.0-1.5) Gyrs between
    snapshots, and the blue lines corresponds to ~(1.0-1.5) Gyrs
    between snapshots. \label{fig:e_spec_tests}}
\end{minipage}
\end{figure*}

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{./figures/CRvolume.g72a.1.4Rv.a24.full.140.v20.eps}
  \includegraphics[width=0.49\columnwidth]{./figures/CRvolume.g1a.1.4Rv.a24.full.020.v20.eps}
  \caption{The fractional volume occupied by CR electrons in cluster
    outskirts. Large panels (left Coma like cluster, and right massive
    cooling flow cluster) show the fractional volume occupied by CR
    electrons as a function of logarithmic Lorentz factor ($\gam_\e$)
    in the region between (1.3-1.5)$\rvir$ at redshift zero. The
    smaller panels show the logarithmic particle distribution function
    for different $\gam_\e$, where the left peaks corresponds to
    unpopulated SPH particles and cooled CR particles, while the right
    peaks show the CR populated SPH
    particles. \label{fig:e_spec_CRvolume}}
\end{minipage}
\end{figure*}

\begin{figure}
  \includegraphics[width=1.0\columnwidth]{./figures/CRespec.all.1.4Rv.a24.full.020.v20.eps}
  \caption{Cosmic ray electron spectra in cluster outskirts at
    redshift zero. We show the total volume weighted CR electron
    distribution function in the region between (1.3-1.5)$\rvir$ with
    full cooling that includes radiative, Coulomb, and adiabatic
    losses. The the thin grey lines show the spectra from individual
    clusters while the thick solid line shows the median of the
    spectra from all clusters. \label{fig:e_spec_all}}
\end{figure}

\begin{figure}
  \includegraphics[width=1.0\columnwidth]{./figures/old/CRespecRadial.Radial2Rv.a24.140.v16.eps}
  \caption{Spatial dependence of the cosmic ray electron spectra at
    redshift zero. We show the total volume weighted CR electron
    distribution function; a Coma like cluster (left panel), a massive
    cooling flow cluster (middle panel), and the median of our 14
    cluster's spectra (right panel). The different line colors show
    the CR electron spectra for different radial
    bins.\label{fig:e_spec_spatial}}
\end{figure}

\begin{figure}
  \includegraphics[width=1.0\columnwidth]{./figures/CRespec.g72a.1.4Rv.a24.full.140.v20.eps}
  \caption{Cosmic ray electron spectra in cluster outskirts at
    redshift zero. We show the CR electron distribution function
    weighted with the electron Lorentz factor ($\gam_\e$) squared for
    a typical cluster with a recent merger for region between
    (1.3-1.5)$\rvir$. The median CR spectrum is shown with full
    cooling that includes radiative, Coulomb, and adiabatic losses
    (dash-dotted line), and with only adiabatic losses (dashed
    line). The solid line represents the total volume weighted CR
    electron spectrum with full cooling. Note that the Coulomb cooling
    at low energies is very inefficient because of the low electron
    densities in the cluster outskirts.\label{fig:e_spec_g72a}}
\end{figure}
